---
title: "GSS challange"
author: "Tattiana Rinkenberger"
date: "07/31/2021"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(knitr)
GSS <-read.csv(file="GSScsv.csv",header=TRUE)
library(dplyr)
GSS <- GSS %>%
mutate(across(where(is.character), as.factor))
```


# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

Is attitude toward capitol punishment (CAPPUN) independent of race (RACE)?
If not, then whose view differs from whom??

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

#### Plot barplots tigerstats way

```{r}
barchartGC(~RACE + CAPPUN,data=GSS)
barchartGC(~RACE + CAPPUN,data=GSS, type="percent")

```

It appears that Black are more likely to be against CAPPUN compared to Other and White who are in favor of CAPPUN.


### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table2 <- xtabs(~RACE + CAPPUN, data=GSS)
rowPerc(table2)
colPerc(table2)
```

Based on the table, Black 34.62 are in favor of the CAPPUN whereas Other is 83.33 and White is 68.66 in favor of CAPPUN. It appears there is a significant difference between Black versus Other and White. 

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisq.test(table2)
print("fisherXY")
chisqtestGC(table2)
```

```{r}
fisherXY(GSS$RACE,GSS$CAPPUN)
```
```{r}
fisherXY(GSS$RACE,GSS$CAPPUN,printALL = TRUE)
```

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

When we looked at the fisherXY test we see that there is a significant difference between Black and Other and White. We also saw that there isnt a big difference between Other and White. 
